---
title: "Nonparametric Homework 3"
author: "Forrest Walker"
output:
  word_document: default
  html_notebook: default
---
## Question 1

### Part A

Compute the observed SST,MST,SSE, and MSE. F=MST/MSE.
```{r}
group_1 <- c(2.9736,0.9448,1.6394,0.0389,1.2958)
group_2 <- c(0.7681,0.8027,0.2156,0.074,1.5076)
group_3 <- c(4.8249,2.2516,1.5609,2.0452,1.0959)
total <- c(group_1,group_2,group_3)

E_g1 <- mean(group_1)
E_g2 <- mean(group_2)
E_g3 <- mean(group_3)
E_total <- mean(total)
obs_SST <- 5*((E_g1-E_total)^2)+(5*((E_g2-E_total)^2))+(5*((E_g3-E_total)^2))
obs_SST
MST <- obs_SST/(3-1)
MST
```

```{r}
S2_1 <- var(group_1)
S2_2 <- var(group_2)
S2_3 <- var(group_3)
SSE <-  (4*S2_1)+(4*S2_2)+(4*S2_3)
MSE <- SSE/(15-3)
obs.F <- MST/MSE 
obs.F 
```
My observed F statistic is 2.990689.  
Now do the permutations,

```{r}
library(gtools)
A <- 10000
Q1_matrix <- matrix(NA,A,15)
Q1_SSTperm <- rep(NA,A)
Q1_MSTperm <- rep(NA,A)
Q1_SSEperm <- rep(NA,A)
Q1_MSEperm <- rep(NA,A)
Q1_Fperm <- rep(NA,A)
Q1_count <- 0
for(i in 1:A) {
  Q1_matrix[i,] <-permute(total)
  Q1_SSTperm[i] <- 5*((mean(Q1_matrix[i,1:5])-mean(Q1_matrix[i,]))^2)
  +(5*((mean(Q1_matrix[i,6:10])-mean(Q1_matrix[i,]))^2))
  +(5*((mean(Q1_matrix[i,11:15])-mean(Q1_matrix[i,]))^2))
  Q1_MSTperm[i] <- Q1_SSTperm[i]/2
  Q1_SSEperm[i] <- 4*(var(Q1_matrix[i,1:5]))+4*(var(Q1_matrix[i,6:10]))+(4*(var(Q1_matrix[i,11:15])))
  Q1_MSEperm[i] <- Q1_SSEperm[i]/12
  Q1_Fperm[i] <- Q1_MSTperm[i]/Q1_MSEperm[i]
  if(Q1_Fperm[i]>=obs.F){
    Q1_count <- Q1_count+1
  }
  }

Q1_count/A
```
 My p value is 0.0059.
 
### Part B

```{r}
rep <- c(rep("g1",5),rep("g2",5),rep("g3",5))
df <- data.frame(total,rep)
huh <- aov(total~rep,data=df)
summary(huh)
```
looks like the one way anova has a higher p value and would fail to reject the null, and the permutation test would reject the null.

## Question 2

First the Anova way,
```{r}
lb_1700 <- c(574,976,789,805,361,529)
lb_2300 <- c(791,1146,394,767,1385,1021,2073,803,1263,1016,1101,945,139)
lb_2800 <- c(865,775,729,1721,1113,820,1613,1404,1201,205,1380,580,1803)
lb_3200 <- c(998,1049,736,782,730,742,1219,705,1260,611,1350,1657,1143)
lb_3700 <- c(1154,541,406,1529,1132,767,1224,314,1728)
Q2_total <- c(lb_1700,lb_2300,lb_2800,lb_3200,lb_3700)
Q2_groups <- c(rep("lb_1700",6),rep("lb_2300",13),rep("lb_2800",13),rep("lb_3200",13),rep("lb_3700",9))
Q2_df <- data.frame(Q2_total,Q2_groups)
Q2_aov <- aov(Q2_total~Q2_groups)
summary(Q2_aov)
```

Then the permutation way, finding the observed F value. 
```{r}
Q2_obs_SST <- 6*((mean(Q2_total[1:6])-mean(Q2_total))^2)+(13*((mean(Q2_total[7:19])-mean(Q2_total))^2))+(13*((mean(Q2_total[20:32])-mean(Q2_total))^2))+(13*((mean(Q2_total[33:45])-mean(Q2_total))^2))+(9*((mean(Q2_total[46:54])-mean(Q2_total))^2))
Q2_obs_MST <- Q2_obs_SST/4
Q2_obs_SSE <- 5*(var(Q2_total[1:6]))+12*(var(Q2_total[7:19]))+(12*(var(Q2_total[19:32])))+(12*(var(Q2_total[32:45])))+(9*(var(Q2_total[45:54])))
Q2_obs_MSE <- Q2_obs_SSE/(54/5)
Q2_obs_F <- Q2_obs_MST/Q2_obs_MSE
Q2_obs_F
```



Finding the p value in the permutation way we get,
```{r}
A <- 10000
Q2_matrix <- matrix(NA,A,length(Q2_total))
Q2_SSTperm <- rep(NA,A)
Q2_MSTperm <- rep(NA,A)
Q2_SSEperm <- rep(NA,A)
Q2_MSEperm <- rep(NA,A)
Q2_Fperm <- rep(NA,A)
Q2_count <- 0
for(i in 1:A) {
  Q2_matrix[i,] <-permute(Q2_total)
  Q2_SSTperm[i] <- 6*((mean(Q2_matrix[i,1:6])-mean(Q2_matrix[i,]))^2)+
    (13*((mean(Q2_matrix[i,7:19])-mean(Q2_matrix[i,]))^2))+
    (13*((mean(Q2_matrix[i,20:32])-mean(Q2_matrix[i,]))^2))+
       (13*((mean(Q2_matrix[i,33:45])-mean(Q2_matrix[i,]))^2))+
          (9*((mean(Q2_matrix[i,46:54])-mean(Q2_matrix[i,]))^2))
  Q2_MSTperm[i] <- Q2_SSTperm[i]/4
  Q2_SSEperm[i] <- 5*(var(Q2_matrix[i,1:6]))+12*(var(Q2_matrix[i,7:19]))+(12*(var(Q2_matrix[i,19:32])))+(12*(var(Q2_matrix[i,32:45])))+(9*(var(Q2_matrix[i,45:54])))
  Q2_MSEperm[i] <- Q2_SSEperm[i]/12
  Q2_Fperm[i] <- Q2_MSTperm[i]/Q2_MSEperm[i]
  if(Q2_Fperm[i]>=Q2_obs_F){
    Q2_count <- Q2_count+1
  }
}
Q2_count <- Q2_count+1
Q2_count/A
```
My P value is 0.5087.



## Question 3

```{r}
Q1_rankedt <- rank(total)
Q1_grp <- c(rep("g1",5),rep("g2",5),rep("g3",5))
Q3_df <- data.frame(Q1_rankedt,Q1_grp)
kruskal.test(Q1_rankedt~Q1_grp,data=Q3_df)
```

## Question 4

```{r}
a <- 10000
Q4_ranked <- c(1,2,3,4,5,6)
Q4_m <- matrix(NA,a,6)
diff <- rep(NA,a)
Q4_kw <- rep(NA,a)

for(i in 1:a){
  Q4_m[i,] <-permute(Q4_ranked)
  diff[i] <- (2*((mean(Q4_m[i,1:2])-7/2)^2))+
  (2*((mean(Q4_m[i,3:4])-7/2)^2))+
  (2*((mean(Q4_m[i,5:6])-7/2)^2))
  Q4_kw[i] <- diff[i]*(12/42)
}
hist(Q4_kw)

```

## Question 5

```{r}
Q5_score <- c(0,2,1,3,1,1,2,0,3,1,3,4,1,3,4,2,2,1,3,4,2,5,3,4)
Q5_grp <- c(rep("g1",6),rep("g2",6),rep("g3",6),rep("g4",6))
Q_5df <- data.frame(Q5_score,Q5_grp)
kruskal.test(Q5_score~Q5_grp,data=Q_5df)
```


## Question 8

```{r}
compact <- c(791,846,1024,1007,399,1279,407,1656,1036,1226)
heavy <- c(423,541,517,1328,471,533,863,786,551,1068)
light <- c(551,1068,757,1114,920,1809,1238,918,1339,603)
medium <- c(712,435,298,733,1200,1701,707,790,800,480)
mpv <- c(1345,1269,1477,758,996,1306,968,943,1026,1564)
pickup <- c(985,1074,742,985,1342,1184,977,1465,892,1074)
van <- c(805,2613,903,949,1183,1051,1387,1320,1434,1603)
Q8_data <-c(compact,heavy,light,medium,mpv,pickup,van)
Q8_l <- length(Q8_data)

mean.vec <- c(mean(compact),mean(heavy),mean(light),mean(medium),mean(mpv),mean(pickup),mean(van))

mean.diff <- abs(outer(mean.vec,mean.vec,"-"))
b <- 5000
Q <- rep(NA,b)
L <- rep(NA,b)

for(i in 1:b){
  index <- sample(1:70)
  T1=Q8_data[index[1:10]]
  T2=Q8_data[index[11:20]]
  T3=Q8_data[index[21:30]]
  T4=Q8_data[index[31:40]]
  T5=Q8_data[index[41:50]]
  T6=Q8_data[index[51:60]]
  T7=Q8_data[index[61:70]]
  m.vec <- c(mean(T1),mean(T2),mean(T3),mean(T4),mean(T5),mean(T6),mean(T7))
  m.diff <- abs(outer(m.vec,m.vec,"-"))
  Q[i] <- max(m.diff)
  L[i] <- m.diff[1,2]
}
quantile(Q,0.9)
quantile(Q,0.95)




```
